---
title: "strings and factors"
author: "Shihui Peng"
date: "2023-10-22"
output: github_document
---

```{r, message=FALSE, echo=FALSE}
library(tidyverse)
library(p8105.datasets)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# strings

## detect

### str_detect()

```{r}
string_vec = c('my', 'name', 'is', 'shihui')

str_detect(string_vec, 'shihui')

str_detect(string_vec, 'm')

str_detect(string_vec, 'S')
```

* to identify the elements of this string that corresponds to my name -- detect a particular pattern with .
  * **`str_detect()`**: given a string or a vector of character obs, detect the presence or absence of a pattern in a string.
    * don't need the whole string to match w the pattern (the 'e' example).
    * uppercase and lowercase are different values when working w strings. (the 'S' example).

### start/end with 

```{r}
string_vec_2 = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec_2, '^i think')

str_detect(string_vec_2, 'i think$')
```

* We can designate matches at the beginning or end of a line.
  * inside the `str_detect()`:
    * **`'^i think'`**: testing strings that start w 'i think' but not end w 'i think'.
    * **`'i think$'`**: testing strings that end w 'i think' rather than begin w 'i think'.

### design a list of char

```{r}
string_vec_3 = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec_3, '[Pp]umpkin')
```

* We can designate a list of characters that will count as a match.
  * **`'[Pp]umpkin'`**: i want both p and P for 'pumpkin' to be detected, so i put P and p into `[]`. the last string is false b/c the remaining characters don't match.

### provide a range of letters/numbers

```{r}
string_vec_4 = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec_4, '[0-9][a-zA-Z]')
```

* We don’t have to list these; instead, you can provide a `range of letters or numbers` that count as a match.
  * **`'[0-9][a-zA-Z]'`**: i want to get a number from 0 to 9 first, followed immediately by a letter from a to z and A to Z
  * if `'^[0-9][a-zA-Z]'`: i want all those above to be at the start of the string.

### `.` matches everything

```{r}
string_vec_5 = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec_5, '7.11')

str_detect(string_vec_5, '7\\.11')
```

* The character `.` matches anything.
  * **`'7.11'`**: the `.` here means i want 7 with another character (can be anything) with 11.
    * the 3rd string does not match b/c there is nothing between 7 and 11 there.
  * **`'7\\.11'`**: what if we just want exactly `'7.11'`? -- need to put 2 slashes before the `.`.

### `\\` for special characters

```{r}
string_vec_6 = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec_6, "\\[")
```

* Some characters are “special”. These include `[` and `]`, `(` and `)`, and `.`. If you want to search for these, you have to indicate they’re special using **`\`**. Unfortunately, `\` is also special, so things get weird.
  * so we need to use `\\`.


## Replace

is there a way to modify that string vector to replace a string or character w another?
```{r}
str_replace(string_vec, 'shihui', 'Shihui Peng')

str_replace(string_vec, 'm', 'Mac')
```
* **`str_replace()`**: replace the string/character w another

# factors
```{r}
fac_sex = factor(c("male", "male", "female", "female"))

fac_sex

as.numeric(fac_sex)
```

* male and female have implemented some rules to determine what level order are. so if convert this to numeric type, 2 and 1 will be distributed

* what happen if i relevel...

```{r}
fac_sex_level = fct_relevel(fac_sex, 'male')

as.numeric(fac_sex_level)
```

* here, i relevel 'male' as the 1st cat, so when converting to numeric type, it shows 1122 instead of 2211 (when female is the 1st cat)
* **`fct_relevel`**: reorder factor levels by hand
* **`fct_recode`**: change factor labels by hand
* **`fct_reorder`**: reorder factors by sorting along another variable
  * lots of factor functions and string functions we can google or search in r!


# NSDUH - string

### import nsduh dara

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) |> 
  html_table() |> 
  first() |>
  slice(-1)
```

### need to tidy this ds

```{r}
data_marj =
  table_marj |> 
  select(-contains('P Value')) |> 
  pivot_longer(
    -State,
    names_to = 'age_year',
    values_to = 'percent'
  ) |> 
  separate(age_year, into = c('age', 'year'), sep = '\\(') |> 
  mutate(
    year = str_replace(year, '\\)', ''),
    percent = str_replace(percent, '[a-c]$', ''),
    percent = as.numeric(percent)
  ) |> 
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

* **`select(-contains('P Value'))`**: remove any column that contains 'P Value'
* **`-State`**: i want to pivot everything that is not in the State col
* **`separate(age_year, into = c('age', 'year'), sep = '\\(')`**
  * separate the col `age_year` into `age` and `year`
  * where to separate? -- `sep =` 
    * separate from the separate character, which is `(` here, and split it apart at this character, and get rid of this separation stuff. but the `(` is a special character, so we need to use `\\` b4 it. -- `'\\('`
* inside the `mutate()`:
  * **``str_replace(year, '\\)', '')`**: in year col, replace the `)` w nothing (nothing inside the '')
* **` filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))`**: filters OUT rows where the "State" col is one of the specified values ("Total U.S.", "Northeast", "Midwest", "South", "West")

* so for `stringr` part and regular expressions in this code:
  * in separate, we split age and year at the open parentheses using `"\\("`
  * we stripped out the close parenthesis in `mutate`
  * to remove character superscripts, we replaced any character using `"[a-c]$"`
 
# NSDUH -- factors 

so now i want to make a plot. Let’s quickly visualize these data for the 12-17 age group; to make the plot readable, we’ll treat State as a factor are reorder according to the median percent value.

```{r}
data_marj |> 
  filter(age == '12-17') |> 
  mutate(
    State = fct_reorder(State, percent)
    ) |> 
  ggplot(aes(x = State, y = percent, color = year)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

* `mutate(State = fct_reorder(State, percent))`: reorder the State col in the order of percent col. Since we have 2 percents for 1 state, r will use the median value of these 2 percents and put state in this order.
  * we can also use sth like `fct_relevel(State, 'Texas', 'New York')` to reorder specific state order.
  * we can use functions from the forcats package (eg.fct_ ), inside tidyverse, to work with character columns as if they were factor variables, without explicitly converting the column to a factor. These functions are designed to provide factor-like behavior for character vectors.
  
* `theme(axis.text.x = element_text(angle = 90, hjust = 1))`: make the lable in x-axis rotate 90 degree

# Weather data

import data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2023-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

## factors matter in making plots

```{r}
weather_df |> 
  mutate(
    name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))
    ) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()

weather_df |> 
  mutate(
    name = fct_reorder(name, tmax)
    ) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin(aes(fill = name), color = 'blue', alpha = 0.5)
```

## factors also matter in statistics
```{r}
weather_df |> 
  lm(tmax ~ name, data = _)
```

* the ordering of factor variables play an important in this case as well. Specifically, the ordering determines the “reference” category, and is something that can be adjusted as needed. here, if we do not define the order, the order will be alphabetic, so the ref grp is CentralPark_NY for above code.

```{r}
weather_df |>
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  lm(tmax ~ name, data = _)
```

* here, we reorder the name, so the ref grp will be Molokai_HI for above code

# Restaurant inspections data

import data and take a look
```{r}
data("rest_inspec")

rest_inspec |> 
  group_by(boro, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(
    names_from = grade, 
    values_from = n
    )
```

maniputate the data
```{r}
rest_inspec_1 =
  rest_inspec |>
  filter(grade %in% c("A", "B", "C"), boro != "Missing") |> 
  mutate(boro = str_to_title(boro))
```

* **`str_to_title()`**: converts text to title case, where the first letter of each word is capitalized.

## str_detect()

```{r}
rest_inspec_1 |> 
  filter(str_detect(dba, '[Pp][Ii][Zz][Zz][Aa]')) |> 
  group_by(boro, grade) |> 
  summarise(n = n()) |> 
  pivot_wider(
    names_from = grade,
    values_from = n
  )
```

* if we use `str_detect(dba, 'pizza')`, it doesn’t look right – for sure there are more pizza place ratings than that! The problem is that the match in str_detect is case-sensitive until we tell it not to be. So instead, we need to use '[Pp][Ii][Zz][Zz][Aa]' inside str_detect().

## fct_infreq()

```{r}
rest_inspec_1 |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(
    boro = fct_infreq(boro)
  ) |> 
  ggplot(aes(x = boro, fill = grade), alpha = 0.3) + 
  geom_bar()
```

* **`fct_infreq()`**: Might help to have things in a different order – maybe number of pizza places in this case. that is use the frequency for different boro as the order in x-axis.

## str_replace()

* Suppose I want to rename a borough. I could try using **`str_replace`**.

```{r}
rest_inspec_1 |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "Manhattan", "The City")) |>
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```

* This renamed the borough, but then converted the result back to a string (`fct_infreq()` does not work out in the plot) – which, when plotted, was implicitly made a factor and ordered alphabetically. I could switch the order in which I rename and encode the factor order I want, but that might not always work.

* but if we reverse the code order into:

```{r}
rest_inspec_1 |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(
    boro = str_replace(boro, "Manhattan", "The City"),
    boro = fct_infreq(boro)
    ) |>
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```
Then the `fct_infreq()` can work.

### if we use `replace()`, sth bad

```{r}
rest_inspec_1 |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(
    boro = fct_infreq(boro),
    boro = replace(boro, which(boro == "Manhattan"), "The City")) |>
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```

* 'NA' is present inthe place where we expect to show 'The City'
  * it replaces the values in the boro column that are equal to "Manhattan" with "The City," but it doesn't modify the levels of the factor. we need to modify 'Manhattan' in factor levels.
  * Factors have very specific values, trying to use a value that is not an existing factor level won’t work.
  
## fct_recode

Fortunately there is a dedicated function for renaming factor levels:
```{r}
rest_inspec_1 |> 
  filter(str_detect(dba, regex("pizza", ignore_case = TRUE))) |>
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "The City" = "Manhattan")) |>
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```

