---
title: "strings and factors"
author: "Shihui Peng"
date: "2023-10-22"
output: github_document
---

```{r, message=FALSE, echo=FALSE}
library(tidyverse)
library(p8105.datasets)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# strings

## detect

### str_detect()

```{r}
string_vec = c('my', 'name', 'is', 'shihui')

str_detect(string_vec, 'shihui')

str_detect(string_vec, 'm')

str_detect(string_vec, 'S')
```

* to identify the elements of this string that corresponds to my name -- detect a particular pattern with .
  * **`str_detect()`**: given a string or a vector of character obs, detect the presence or absence of a pattern in a string.
    * don't need the whole string to match w the pattern (the 'e' example).
    * uppercase and lowercase are different values when working w strings. (the 'S' example).

### start/end with 

```{r}
string_vec_2 = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec_2, '^i think')

str_detect(string_vec_2, 'i think$')
```

* We can designate matches at the beginning or end of a line.
  * inside the `str_detect()`:
    * **`'^i think'`**: testing strings that start w 'i think' but not end w 'i think'.
    * **`'i think$'`**: testing strings that end w 'i think' rather than begin w 'i think'.

### design a list of char

```{r}
string_vec_3 = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec_3, '[Pp]umpkin')
```

* We can designate a list of characters that will count as a match.
  * **`'[Pp]umpkin'`**: i want both p and P for 'pumpkin' to be detected, so i put P and p into `[]`. the last string is false b/c the remaining characters don't match.

### provide a range of letters/numbers

```{r}
string_vec_4 = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec_4, '[0-9][a-zA-Z]')
```

* We don’t have to list these; instead, you can provide a `range of letters or numbers` that count as a match.
  * **`'[0-9][a-zA-Z]'`**: i want to get a number from 0 to 9 first, followed immediately by a letter from a to z and A to Z
  * if `'^[0-9][a-zA-Z]'`: i want all those above to be at the start of the string.

### `.` matches everything

```{r}
string_vec_5 = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec_5, '7.11')

str_detect(string_vec_5, '7\\.11')
```

* The character `.` matches anything.
  * **`'7.11'`**: the `.` here means i want 7 with another character (can be anything) with 11.
    * the 3rd string does not match b/c there is nothing between 7 and 11 there.
  * **`'7\\.11'`**: what if we just want exactly `'7.11'`? -- need to put 2 slashes before the `.`.

### `\\` for special characters

```{r}
string_vec_6 = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec_6, "\\[")
```

* Some characters are “special”. These include `[` and `]`, `(` and `)`, and `.`. If you want to search for these, you have to indicate they’re special using **`\`**. Unfortunately, `\` is also special, so things get weird.
  * so we need to use `\\`.


## Replace

is there a way to modify that string vector to replace a string or character w another?
```{r}
str_replace(string_vec, 'shihui', 'Shihui Peng')

str_replace(string_vec, 'm', 'Mac')
```
* **`str_replace()`**: replace the string/character w another

# factors
```{r}
fac_sex = factor(c("male", "male", "female", "female"))

fac_sex

as.numeric(fac_sex)
```

* male and female have implemented some rules to determine what level order are. so if convert this to numeric type, 2 and 1 will be distributed

* what happen if i relevel...

```{r}
fac_sex_level = fct_relevel(fac_sex, 'male')

as.numeric(fac_sex_level)
```

* here, i relevel 'male' as the 1st cat, so when converting to numeric type, it shows 1122 instead of 2211 (when female is the 1st cat)
* **`fct_relevel`**: reorder factor levels by hand
* **`fct_recode`**: change factor labels by hand
* **`fct_reorder`**: reorder factors by sorting along another variable
  * lots of factor functions and string functions we can google or search in r!
